---
title: "Scalable and efficient functional connectivity modelling"
author: "XXXXXXXXXXXXXXXXXXXX"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
---

This script contains the R pipeline for tiling and buffering the landscape, which is steps 1 and 2 of our tiling approach.



# libraries

```{r, warning=FALSE, message = FALSE, error = FALSE}
library(tidyverse)
library(samc)
library(terra)
library(tidyterra)
library(raster)
library(parallel)
library(pryr)
library(doParallel)
library(foreach)
```

# Upload case study information

```{r}
caseStudy <- readRDS("../data/107a_1000_99/107a_1000_99.rds")
caseStudy
```

# Upload the landscape
- resistance raster
- source raster
- absorption raster

```{r}
samc_stack <- rast("../data/107a_1000_99/107a_1000_99_stack.tif")
plot(samc_stack)
```


# Loading the tiling function

```{r}
source("../functions/SAMC_tiling.R")
```

THE FUNCTION:

samc_tiling(dataResistance, # Resistance raster
            dataAbsorption, # Absorption raster
            dataInitiation, # Initiation raster
            extentFocalArea, # Extent of focal area
            tileDimensions, # 2 values for number of rows and cols for the zoning of tiles
            bufferDistance, # Buffer distance
            purgeExistingTiles = TRUE, # Whether to delete previous files of same modelName
            modelname)


Tiling (i.e., spatial chunking) the modelling domain involves two processes: 
(1) make the "zones" or the tiling blueprint,
(2) then tile (i.e., spatially slice) the rasters using the zones.

OUTPUT is a list containing filepaths for the 

```{r}
### tiles ###
tile_zone <- rast(
                  nrows= 5, # number of tile rows
                  ncols = 5, # number of tile cols
                  extent = caseStudy$focal_extent # Extent of main landscape
                  )
plot(samc_stack$init, main = "Tiling zones (25 tiles)")
plot(as.polygons(tile_zone, dissolve = FALSE), color = "white", add = TRUE)
```


# Tiling the raster

```{r}
for (i in c(2,3,4,5)) {
  data_tiles <- samc_tiling(domainName = caseStudy$landscapeName,
                            dataResistance = samc_stack$res,
                            dataAbsorption = samc_stack$abs,
                            dataInitiation = samc_stack$init,
                            extentFocalArea = caseStudy$focal_extent,
                            tileDimensions = c(i,i),
                            bufferDistance = caseStudy$buffer, 
                            purgeExistingTiles = TRUE)
  
}

par(mfrow = c(1,3))
plot(rast(data_tiles$resistance[1]), main = "Resistance tile\n(1 of 25)")
plot(rast(data_tiles$initiation[1]), main = "Source tile\n(1 of 25)")
plot(rast(data_tiles$absorption[1]), main = "Absorption tile\n(1 of 25)")
```

Notice the buffer region of the Source tile; cell values are set to zero so that no movement starts at he buffers.

```{r}
library(grateful)
grateful::cite_packages(output = "file", out.dir = "../", out.format = "docx", out.file = "references_tilingAndBuffering, skip.missing = TRUE)
```

