---
title: "Scalable and efficient functional connectivity modelling"
author: "XXXXXXXXXXXXXXXXXXXX"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
---

This script contains the R pipeline for modelling connectivity on each buffered tile and merging the outputs of buffered tiles, which correspond to steps 3 and 4 of our tiling approach.


# libraries

```{r, warning=FALSE, message = FALSE, error = FALSE}
library(tidyverse)
library(samc)
library(terra)
library(tidyterra)
library(raster)
library(parallel)
library(pryr)
library(doParallel)
library(foreach)
```

# Upload case study information

```{r}
caseStudy <- readRDS("../data/107a_1000_99/107a_1000_99.rds")
caseStudy
```

# The set of tiles you want to model
```{r}
nTiles <- 9
```

# Loading the function

```{r}
source("../functions/SAMC_benchmarking.R")
```

THE FUNCTION:

samc_tiling(
- *dataResistance*, # one .tif file OR vector string of pathnames to several Resistance raster tiles
- *dataAbsorption*, # one .tif file OR vector string of pathnames to several Absorption raster tiles
- *dataInitiation*, # one .tif file OR vector string of pathnames to several Source raster tiles
- *transitionModel*, # list containing arguments on how transition matrix will be calculated from the provided input rasters. Default: list(fun = function(x) 1/mean(x), dir = 8, sym = TRUE)
- *analysis*, # can be one or several SAMC metrics: c("distribution", "mortality", "survival", "visitation", "visitation_net"). Default: all mentioned metrics
- *domainExtent* # extent of the modelling domain (the extent of the original landscape, focal area + buffer)
- *domainName*, # naming convention of the domain (for making folders)
- *experimentName*, # the experiment name (for making folders)
- *nCores*, # the numeber of processors when running in parallel; Default: 1 (no parallelisation)
- *replicateNumber*, # number of replicates Default: 1
- *extentFocalArea*, # Extent of focal area
- *saveRaster*, #  whether the output rasters should be saved. Default: TRUE
- *purgeExistingTiles*, # Whether to delete previous files of same modelName. Default: FALSE
- *outputFolder*, #directory where outputs will be saved default: "../output",
- *saveMeta*) # whether metadata should be saved. Default: TRUE,


# SAMC connectivity modelling and merging

```{r}
data_tiles <- readRDS(paste0("../input/107a_1000_99/", 
                               "tiles_",nTiles, "/",
                               "107a_1000_99_tiles_",nTiles, ".rds"))
  
samc_out <- samc_benchmarking(dataResistance = data_tiles$resistance,
                                dataAbsorption = data_tiles$absorption,
                                dataInitiation = data_tiles$initiation,
                                analysis = c("visitation", "mortality"),
                                domainExtent = ext(data_tiles$domainExtent),
                                domainName  = caseStudy$landscapeName,
                                experimentName = "107a_1000_99_9",
                                nCores = 1, 
                                replicateNumber = 1,
                                saveRaster = TRUE)
```

```{r}
par(mfrow = c(1,2))
plot(rast(samc_out$analysisOutput$visitation), main = "Visitation")
plot(rast(samc_out$analysisOutput$mortality), main = "mortality")
```


```{r}
library(grateful)
grateful::cite_packages(output = "file", 
                        out.dir = "../", 
                        out.format = "docx",
                        out.file = "references_SAMCandMerging", 
                        skip.missing = TRUE)
```

