---
title: "Scalable and efficient functional connectivity modelling"
author: "XXXXXXXXXXXXXXXXXXXX"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
---

This script contains the R pipeline for the sensitivity analysis of spatial accuracy due to tiling and buffering.

# Upload libraries

```{r, warning=FALSE, message = FALSE, error = FALSE}
library(tidyverse)
library(samc)
library(terra)
library(tidyterra)
library(khroma)
library(raster)
library(SSIMmap)
library(lme4)
library(car)
library(broom)
library(MuMIn)
library(arm)
library(glmmTMB)
```


# Upload metadata of the landscape

```{r}
caseStudy <- readRDS("../data/cs_107a_1000_99.rds")
```

# Upload the reference (no tiling) and merged (with tiling) connectivity maps.


```{r}
reference_raster <- rast(paste0("../output/107_a_1000_99/107_a_1000_99_1/visitation/visitation_1.tif"))

tiledOutput_raster <- rast(reference_raster)

for (i in c("4", "9", "16", "25")) {
  tiledOutput_raster <- c(tiledOutput_raster,
                          rast(paste0("../output/107_a_1000_99/107_a_1000_99_",
                                      i, "/visitation/visitation_1.tif"
                                      )
                               ))
}
names(tiledOutput_raster) <- c("4", "9", "16", "25")

reference_raster <- crop(reference_raster,
                         ext(caseStudy$focal_extent))

tiledOutput_raster <- crop(tiledOutput_raster,
                         ext(caseStudy$focal_extent))



plotdata <- as.data.frame(tiledOutput_raster, xy = TRUE) %>% 
  pivot_longer(cols = c("4", "9", "16", "25"), 
               names_to = "tiles", 
               values_to = "visitation") %>% 
  mutate(tiles = factor(tiles,
                        levels = c("4", "9", "16", "25")))

plotdata %>% 
  ggplot() +
  geom_raster(aes(x = x,
                  y = y,
                  fill = visitation)) + 
  facet_grid(cols = vars(tiles)) +
  scale_fill_bamako(reverse = TRUE, 
                    transform = "sqrt", 
                    na.value = NA) + 
  theme_void() +
  theme(aspect.ratio = 1,
        legend.key.width = unit(0.25, 'cm'),
        legend.key.height = unit(0.5, 'cm'))
              
```


```{r}
ggsave(plot = last_plot(), 
       filename = "../figures/SA_visitation.png", 
       dpi = 300, 
       units = "mm", 
       height = 40, 
       width = 200)
```


# Compute difference of reference (no tiling) and merged (with tiling) connectivity maps.


```{r}
error_raster <- rast(reference_raster)

tiles <- c("4", "9", "16", "25")
for (i in 1:length(tiles)) {
  error_raster <- c(error_raster,
                  tiledOutput_raster[[i]] - reference_raster)
}


plot_range <- c(min(minmax(error_raster)[1,]),
                max(minmax(error_raster)[2,]))


plotdata <- as.data.frame(error_raster, xy = TRUE) %>% 
  pivot_longer(cols = c("4", "9", "16", "25"), 
               names_to = "tiles", 
               values_to = "difference") %>% 
  mutate(tiles = factor(tiles,
                        levels = c("4", "9", "16", "25")))

plotdata %>% 
  ggplot() +
  geom_raster(aes(x = x,
                  y = y,
                  fill = difference)) + 
  facet_grid(cols = vars(tiles)) +
  scale_fill_bam() + 
  theme_void() +
  theme(aspect.ratio = 1,
        legend.key.width = unit(0.25, 'cm'),
        legend.key.height = unit(0.5, 'cm'))
              
```


```{r}
ggsave(plot = last_plot(), 
       filename = "../figures/SA_difference.png", 
       dpi = 300, 
       units = "mm", 
       height = 40, 
       width = 200)
```


# Compute spatial similarity of reference (no tiling) and merged (with tiling) connectivity maps.

```{r}
normalize <- function(x) {
  (x - min(x[], na.rm=TRUE)) / (max(x[], na.rm=TRUE) - min(x[], na.rm=TRUE))
}

reference_img <- app(reference_raster, normalize)

tileOutput_img <- rast(tiledOutput_raster)

for (i in 1:length(names(tiledOutput_raster))) {
  tileOutput_img <- c(tileOutput_img, 
                      app(tiledOutput_raster[[i]], normalize)
                    )

}
tiles <- c("4", "9", "16", "25")

names(tileOutput_img) <- tiles


SSIM_raster <- rast(reference_raster)

for (i in 1:length(tiles)) {
  SSIM_raster <- c(SSIM_raster, 
                   ssim_raster(reference_img, 
                               tileOutput_img[[i]], 
                               global = FALSE)[[1]]
  )
  
}
names(SSIM_raster) <- tiles
```


```{r}
plotdata <- as.data.frame(SSIM_raster, xy = TRUE) %>% 
  pivot_longer(cols = c("4", "9", "16", "25"), 
               names_to = "tiles", 
               values_to = "SSIM") %>% 
  mutate(tiles = factor(tiles,
                        levels = c("4", "9", "16", "25")))

plotdata %>% 
  ggplot() +
  geom_raster(aes(x = x,
                  y = y,
                  fill = SSIM)) + 
  facet_grid(cols = vars(tiles)) +
  scale_fill_bam() + 
  theme_void() +
  theme(aspect.ratio = 1,
        legend.key.width = unit(0.25, 'cm'),
        legend.key.height = unit(0.5, 'cm'))
              
```


```{r}
ggsave(plot = last_plot(), 
       filename = "../figures/SA_ssim.png", 
       dpi = 300, 
       units = "mm", 
       height = 40, 
       width = 200)
```


# Compute global spatial error metrcs

- *MAD* - Mean Absolute Difference
- *RMSD* - Root Mean Square Difference
- *SSIM* - Spatial Similarit


```{r}
bind_cols(data.frame(metrics = c("MAD", "RMSD", "SSIM")),
          bind_rows(
as.data.frame(error_raster, xy = TRUE) %>% 
  dplyr::select(-c(x,y)) %>% 
  summarise(across(everything(), ~ mean(abs(.x), na.rm = TRUE))),

as.data.frame(error_raster, xy = TRUE) %>% 
  dplyr::select(-c(x,y)) %>% 
  summarise(across(everything(), ~ round(sqrt(mean(.x^2, na.rm = TRUE)), 4))),

as.data.frame(SSIM_raster, xy = TRUE) %>% 
  dplyr::select(-c(x,y)) %>% 
  summarise(across(everything(), ~ mean(.x, na.rm = TRUE)))
)
)
```



## All the data

```{r}
head(SA_Results_all <- read.csv("../output/SA_Results_107.csv"))
```



```{r}
SA_Results_all %>% 
  mutate(T_bufferDist = factor(T_bufferDist)) %>% 
  ggplot(aes(x = T_nTiles,
             y = value,
             color = T_bufferDist,
             group = T_bufferDist)) +
  geom_point(alpha = 0.2) +
  stat_summary(geom = "pointrange", fun.data = "mean_cl_boot", size = .5) +
  stat_summary(geom = "line", fun = "mean") +
#  guides(color = "none") +
  labs(y = "",
       x = "# of Tiles",
       color = "Dispersal distance") +
  facet_grid(rows = vars(metric), 
             cols = vars(buffer),
             scale = "free" ,
             switch="y") +
  scale_x_continuous(breaks = c(4, 9, 16, 25)) +
  scale_color_bamako(discrete = TRUE, 
                     reverse = TRUE) +
#  scale_color_muted() +
  theme_bw() +
  theme(legend.position = "bottom",
        strip.background = element_blank(),
        strip.placement = "outside",
        aspect.ratio = 1)
```


```{r}
ggsave(plot = last_plot(), 
       filename = "../figures/sensitivityAnalysis.png", 
       dpi = 300, 
       units = "mm", 
       width = 200)
```



# Statistical test

```{r}
head(glmm_data <- SA_Results_all %>% 
  rename("tiles" = "T_nTiles",
         "landscape" = "focalAreaName") %>% 
    na.omit()
)
```

## Fixed effects of tile number and buffer size on RMSD

```{r, fig.width=12}
glm_rmsd <- glmer(data = glmm_data %>%  filter(metric == "RMSD"), 
                         formula = value ~ tiles*buffer + (1|T_bufferDist)  + (1|landscape), 
                         family = Gamma(link = "log"))
summary(glm_rmsd)
r.squaredGLMM(glm_rmsd)
car::Anova(glm_rmsd)

# diagnostic plot
a <- plot(glm_rmsd, 
          type=c("p","smooth"), 
          col.line=1,
          main = "fitted vs residual with smooth line")
b <- plot(glm_rmsd,
          sqrt(abs(resid(.)))~fitted(.),
          type=c("p","smooth"), 
          col.line=1,
          main = "scale-location")
c <- lattice::qqmath(glm_rmsd, 
                     main = "Q-Q")
d <- plot(glm_rmsd, rstudent(.) ~ hatvalues(.), 
          main = "residuals vs leverage")
cowplot::plot_grid(a,b,c,d)
```

## Fixed effects of tile number and buffer size on MAD

```{r, fig.width=12}
glm_mad<- glmer(data = glmm_data %>%  filter(metric == "MAD"), 
                         formula = value ~ tiles*buffer + (1|T_bufferDist) + (1|landscape), 
                         family = Gamma(link = "log"))
summary(glm_mad)
r.squaredGLMM(glm_mad)
car::Anova(glm_mad)

# diagnostic plot
a <- plot(glm_mad, 
          type=c("p","smooth"), 
          col.line=1,
          main = "fitted vs residual with smooth line")
b <- plot(glm_mad,
          sqrt(abs(resid(.)))~fitted(.),
          type=c("p","smooth"), 
          col.line=1,
          main = "scale-location")
c <- lattice::qqmath(glm_mad, 
                     main = "Q-Q")
d <- plot(glm_mad, rstudent(.) ~ hatvalues(.), 
          main = "residuals vs leverage")
cowplot::plot_grid(a,b,c,d)
```

## Fixed effects of tile number and buffer size on SSIM


```{r, fig.width=12}
glm_ssim<- glmmTMB(data = glmm_data %>%  filter(metric == "SSIM"), 
                         formula = value ~ tiles*buffer + (1| T_bufferDist) + (1|landscape), 
                         family = beta_family(link = "logit"))
summary(glm_ssim)
r.squaredGLMM(glm_ssim)
car::Anova(glm_ssim)

# diagnostic plot
binnedplot(predict(glm_ssim,type='response'),
           resid(glm_ssim,type='response'))
```

# References

```{r}
library(grateful)
grateful::cite_packages(output = "file", out.dir = "../", out.format = "docx", out.file = "references_sensitivityAnalysis", skip.missing = TRUE)
```

