---
title: "Scalable and efficient functional connectivity modelling"
author: "XXXXXXXXXXXXXXXXXXXX"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
---

This script contains the R pipeline for the benchmarking of runtime and maximum memory usage.

# Upload libraries
```{r, warning=FALSE, message = FALSE, error = FALSE}
library(tidyverse)
library(cowplot)
library(khroma)
library(terra)
library(lme4)
library(car)
library(broom)
library(MuMIn)
library(lattice)
```

# Upload the benchmarking data

```{r}
head(benchmark_data <- read.csv("../output/Benchmarking_Results_107.csv")
)
```

# Benchmarking results

```{r fig.width=3, fig.height=6}
a <- benchmark_data %>%
  mutate(nCores = factor(nCores)) %>% 
  ggplot(aes(y = runTime, 
             x = T_nTiles)) +
  geom_point(aes(color = nCores,
                   group = nCores),
             position = position_jitter(width = 0.2),
             alpha = 0.2, 
             size = 1) +
  stat_summary(aes(color = nCores,
                   group = nCores),
               geom = "point", 
               size = 2) +
  stat_summary(aes(color = nCores,
                   group = nCores),
               geom = "line",
               size = 1) +
  labs(y = "Runtime (H)",
       x = "# of Tiles",
       color = "# of Cores") +
  scale_x_continuous(breaks = c(1, 4, 9, 16, 25)) +
  scale_color_bamako(reverse = TRUE, 
                    range = c(0.2,1),
                    breaks = c(1,2,4,6,8,10),
                    discrete = TRUE) +
  theme_bw() +
  theme(legend.position = "none")+
  theme(aspect.ratio = 1)

b <- benchmark_data %>% 
  mutate(nCores = factor(nCores)) %>% 
  ggplot(aes(y = runTime_speedup, 
             x = T_nTiles)) +
  geom_point(aes(color = nCores,
                   group = nCores),
             position = position_jitter(width = 0.2),
             alpha = 0.2, size = 1) +
  stat_summary(aes(color = nCores,
                   group = nCores),
               geom = "point",
               size = 2) +
  stat_summary(aes(color = nCores,
                   group = nCores),
               geom = "line",
               size = 1) +
  labs(y = "Speed up",
       x = "# of Tiles",
       color = "# of Cores") +
  scale_x_continuous(breaks = c(1, 4, 9, 16, 25)) +
  scale_color_bamako(reverse = TRUE, 
                    range = c(0.2,1),
                    breaks = c(1, 2,4,6,8,10),
                    discrete = TRUE) +
  theme_bw()+
  theme(legend.position = "none")+
  theme(aspect.ratio = 1)

c <- benchmark_data %>% 
  mutate(nCores = factor(nCores)) %>% 
  ggplot(aes(y = memory, 
             x = T_nTiles)) +
  geom_point(aes(color = nCores,
                   group = nCores),
             position = position_jitter(width = 0.2),
             alpha = 0.2, 
             size = 1) +
  stat_summary(aes(color = nCores,
                   group = nCores),
               geom = "point",
               size = 2) +
  stat_summary(aes(color = nCores,
                   group = nCores),
               geom = "line",
               size = 1) +
  labs(y = "Max Memory Usage (GB)",
       x = "# of Tiles",
       color = "# of Cores") +
  guides(color = guide_legend(nrow = 2)) +
  scale_x_continuous(breaks = c(1, 4, 9, 16, 25)) +
  scale_color_bamako(reverse = TRUE, 
                    range = c(0.2,1),
                    breaks = c(1,2,4,6,8,10),
                    discrete = TRUE) +
  theme_bw()+
  theme(legend.position = "bottom",
        aspect.ratio = 1)

plot_grid(a,b,c, nrow = 3, rel_heights =  c(1,1,1.4))
```


```{r}
ggsave(plot = last_plot(), 
       filename = "../figures/benchmark.jpg", 
       dpi = 300, 
       units = "mm",  
       width = 90)
```


# Statistical test

```{r}
glmm_data <- benchmark_data %>%
  mutate(T_nTiles = scale(T_nTiles),
         nCores = scale(nCores)) %>% 
  rename("tiles" = "T_nTiles",
         "cores" = "nCores",
         "landscape" = "focalArea")
```


## Fixed effects of tile number and parallelisation on runtime




```{r, fig.width=12}
glm_runTime <- glm(data = glmm_data, 
                   formula = runTime ~ tiles*cores + landscape, 
                   family = Gamma(link = "log"))
summary(glm_runTime)
r.squaredGLMM(glm_runTime)
car::Anova(glm_runTime)


tidy(glm_runTime) %>% 
  write.csv("../output/glm_summary_runTime.csv")

tidy(car::Anova(glm_runTime))%>% 
  write.csv("../output/glm_anova_runTime.csv")


# diagnostic plot
plot(glm_runTime)
```

### Permutation importance

```{r}
R_sq <- rep(NA, 4)
AIC_delta <- rep(NA,4)

glm_null <- glm(data = glmm_data, 
                   formula = runTime ~ landscape, 
                   family = Gamma(link = "log"))
R_sq[1] <- r.squaredGLMM(glm_null)[2,2]
AIC_delta[1] <- AICc(glm_null)

glm_1 <- glm(data = glmm_data, 
                   formula = runTime ~ tiles + landscape, 
                   family = Gamma(link = "log"))
R_sq[2] <- r.squaredGLMM(glm_1)[2,2]
AIC_delta[2] <- AICc(glm_1)

glm_2 <- glm(data = glmm_data, 
                   formula = runTime ~ cores + landscape, 
                   family = Gamma(link = "log"))
R_sq[3] <- r.squaredGLMM(glm_2)[2,2]
AIC_delta[3] <- AICc(glm_2)

glm_3 <- glm(data = glmm_data, 
                   formula = runTime ~ tiles + cores + landscape, 
                   family = Gamma(link = "log"))
R_sq[4] <- r.squaredGLMM(glm_3)[2,2]
AIC_delta[4] <- AICc(glm_3)

glm_4 <- glm(data = glmm_data, 
                   formula = runTime ~ tiles*cores + landscape, 
                   family = Gamma(link = "log"))
R_sq[5] <- r.squaredGLMM(glm_4)[2,2]
AIC_delta[5] <- AICc(glm_4)

data.frame(glm = c("null", "tiles", "core", "tiles + core", "tiles * core"),
           R2 = R_sq,
           AICc = min(AIC_delta)- AIC_delta)

data.frame(glm = c("null", "tiles", "core", "tiles + core", "tiles * core"),
           R2 = R_sq,
           AICc = AIC_delta-min(AIC_delta)) %>% 
  write.csv("../output/glm_runtime.csv")


anova(glm_null, glm_1, glm_2, glm_3, glm_4)

```




## Fixed effects of tile number and parallelisation on maximum memory usage

```{r, fig.width=12}
glm_memory <- glm(data = glmm_data,
                  formula = memory ~ tiles*cores + landscape, 
                  family = Gamma(link = "log"))
summary(glm_memory)
r.squaredGLMM(glm_memory)
car::Anova(glm_memory)

tidy(glm_memory) %>% 
  write.csv("../output/glm_summary_memory.csv")

tidy(car::Anova(glm_memory))%>% 
  write.csv("../output/glm_anova_memory.csv")


# diagnostic plot
plot(glm_runTime)
```






### Permutation importance

```{r}
R_sq <- rep(NA, 4)
AIC_delta <- rep(NA,4)

glm_null <- glm(data = glmm_data, 
                   formula = memory ~ landscape, 
                   family = Gamma(link = "log"))
R_sq[1] <- r.squaredGLMM(glm_null)[2,2]
AIC_delta[1] <- AICc(glm_null)

glm_1 <- glm(data = glmm_data, 
                   formula = memory ~ tiles + landscape, 
                   family = Gamma(link = "log"))
R_sq[2] <- r.squaredGLMM(glm_1)[2,2]
AIC_delta[2] <- AICc(glm_1)

glm_2 <- glm(data = glmm_data, 
                   formula = memory ~ cores + landscape, 
                   family = Gamma(link = "log"))
R_sq[3] <- r.squaredGLMM(glm_2)[2,2]
AIC_delta[3] <- AICc(glm_2)

glm_3 <- glm(data = glmm_data, 
                   formula = memory ~ tiles + cores + landscape, 
                   family = Gamma(link = "log"))
R_sq[4] <- r.squaredGLMM(glm_3)[2,2]
AIC_delta[4] <- AICc(glm_3)

glm_4 <- glm(data = glmm_data, 
                   formula = memory ~ tiles*cores + landscape, 
                   family = Gamma(link = "log"))
R_sq[5] <- r.squaredGLMM(glm_4)[2,2]
AIC_delta[5] <- AICc(glm_4)

data.frame(glm = c("null", "tiles", "core", "tiles + core", "tiles * core"),
           R2 = R_sq,
           AICc = min(AIC_delta)- AIC_delta)

data.frame(glm = c("null", "tiles", "core", "tiles + core", "tiles * core"),
           R2 = R_sq,
           AICc = AIC_delta-min(AIC_delta)) %>% 
  write.csv("../output/glm_memory.csv")


anova(glm_null, glm_1, glm_2, glm_3, glm_4)

```



# Reference

```{r}
library(grateful)
grateful::cite_packages(output = "file", out.dir = "../", out.format = "docx", out.file = "references_benchmarking", skip.missing = TRUE)
```

