---
title: "Scalable and efficient functional connectivity modelling"
author: "XXXXXXXXXXXXXXXXXXXX"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
---

This script contains the R pipeline for creating modelling landscapes. For demonstration, we showed here how to create neutral landscape (10 million grid cells).



# libraries

```{r, warning=FALSE, message = FALSE, error = FALSE}
library(tidyverse)
library(samc)
library(terra)
library(tidyterra)
library(raster)
library(gstat)
library(NLMR)
library(landscapetools)
```

# Creating a neutral landscape model

Neutral landscape models were created using a "mid-point displacement approach" implemented in NLMR.

Key parameters of the landscape model:
- roughness = 0.9



## Writing metadata 

Key parameters:
- Focal area size = 10 million cells
- Focal area (equivalent) resolution = 100-m
- Mean dispersal distance in m = 1000
- Buffer percentile = 0.99

All other parameters will be computed from this information



```{r}
caseStudy <- list(
                  landscapeName = NA, #the name of the landscape.
                  resolution = NA, # the resolution of the landscape, useful when buffer is in m
                  displacement = NA, # the mean dispersal distance of the species
                  buffer_quartile = NA, # the quartile/percenfocal used in calculating the buffer size.
                  buffer_in_m = NA, # the raw buffer, when based on 95th percentile or something
                  buffer = NA,  # the number of pixels of the size of the buffer. 
                  domain_extent = NA, # the extent of the entire modelling domain (focal + buffer)
                  focal_cellNo = NA, # the number of cells in the focal domain
                  focal_rows = NA, # the number of rows in the focal region, or sqrt(focal_cellNo if square)
                  focal_columns = NA, # the number of columns in the focal region, or sqrt(focal_cellNo if square)
                  focal_extent = NA # the extend of the focal region
                  )

caseStudy$focal_cellNo <- 10000000

caseStudy$focal_rows <- round(sqrt(caseStudy$focal_cellNo), digits = -3)

caseStudy$focal_columns <- round(sqrt(caseStudy$focal_cellNo), digits = -3)

caseStudy$resolution <- 100

caseStudy$displacement <- 1000

caseStudy$buffer_quartile <- 0.99



caseStudy$buffer_in_m <- -log(1 - caseStudy$buffer_quartile)/(1/caseStudy$displacement)

caseStudy$buffer <- ceiling(caseStudy$buffer_in_m/caseStudy$resolution)

caseStudy$domain_extent <- c(
                             0, # x start
                            (caseStudy$focal_rows + (2*caseStudy$buffer)), #x end
                             0, #y start
                            (caseStudy$focal_columns + (2*caseStudy$buffer)) #y end
                             )

caseStudy$focal_extent <- c(
                         caseStudy$buffer, # x start
                         (caseStudy$buffer + caseStudy$focal_rows), #x end
                         caseStudy$buffer, #y start
                         (caseStudy$buffer + caseStudy$focal_columns) #y end
                         )

caseStudy$landscapeName <- "107a_1000_99"

caseStudy

dir.create(paste0("../data/", caseStudy$landscapeName), 
           recursive = TRUE)
saveRDS(caseStudy,
        file = paste0("../data/", caseStudy$landscapeName, "/",
                      caseStudy$landscapeName,
                      ".rds"))

```

## caststudy rasters

```{r}
# rasterising the modelling domain
domain_rast <- rast(ext = ext(caseStudy$domain_extent), res = 1, vals = 1)

# rasterising the focal region
focal_rast <- rast(ext = ext(caseStudy$focal_extent), res = 1, vals = 1)

# rasterising the outer buffer
outerBuffer <- domain_rast - extend(focal_rast, ext(domain_rast), fill = 0)
outerBuffer[outerBuffer == 0] <- NA

plot(domain_rast)
plot(focal_rast, add = TRUE, col = "green")
```
Map legend:
Green = focal area (area where we want to model connectivity)
Yellow = outer buffers.

```{r}
set.seed(49)
nlm_out <- nlm_mpd(ncol = caseStudy$domain_extent[2],
                       nrow = caseStudy$domain_extent[4],
                       resolution = 1,
                       roughness = 0.80)
plot(nlm_out, main = "Neutral Landscape Model (roughness = 0.8, seed = 49)")
```


### Source raster

Key parameters:
Proportion of habitat (vs matrix) = 0.3

```{r}
init_r <- util_classify(x = nlm_out, 
                        weighting = c(0.7, 0.3))
init_r <- init_r - 1
init_r <- extend(x = rast(init_r), 
               y = ext(caseStudy$domain_extent),
               fill = 0)
plot(init_r, main = "Source raster (30% source habitat)")
```
Map legend:
Violet = matrix
Yellow = habitat

### Resistance raster

```{r}
set.seed(49)
res_r <- 1 - rast(nlm_out)

#re-scale resistance to 1-100
res_r <- stretch(res_r, 
                 minv = 1, 
                 maxv = 100)
res_r <- extend(x = res_r, 
               y = ext(caseStudy$domain_extent),
               fill = 100)

#adjust resistance in habitat
res_r[init_r != 0] <- 1
plot(res_r, main = "Resistance raster")
```


### Absorption raster

```{r}
pixel_dist <- caseStudy$displacement/caseStudy$resolution
abs <- exp(-1.247 - 1.788*log(pixel_dist)) #value for dispersal from Fletcher et al. 2023
abs_r <- rast(res_r, vals = abs)
plot(abs_r, main = "Absorption raster")
```

## Saving the Landscape
```{r}
samc_stack <- c(res_r, abs_r, init_r)
names(samc_stack) <- c("res", "abs", "init")
samc_stack[any(is.na(samc_stack))] <- NA



writeRaster(samc_stack, paste0("../data/",
                               caseStudy$landscapeName, "/",
                               caseStudy$landscapeName, "_stack.tif"), overwrite=TRUE)



#plot samc_stack
plot(samc_stack)
```
```{r}
library(grateful)
grateful::cite_packages(output = "file", out.dir = "../", out.format = "docx", out.file = "references_dataPreparation", skip.missing = TRUE)
```

